Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-06-05 SS3_Krill_481.Rmd
adding html but not in page
test model to example

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
```

@@

The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity. The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill. By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics. This nuanced approach allows for a more refined understanding of different areas features krill structure, fishing mortality and recruitment patterns across the Antarctic Peninsula. Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

*Keywords: Krill populations, dynamic population, stock assessment, SS3, spatial implicit, management, CCAMLR.*

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR routinely conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. This regional fisheries management organization has intensified its efforts to promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].

Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill dynamics, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data on krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.

In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach that incorporates the fishing and biological history of krill over the past 20 years, as well as ecosystem variables such as environmental conditions, primary productivity levels affecting krill recruitment, and predator abundance. Unlike previous assessments, this apporach integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics. Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.

Beyond integrating available information for assessing krill population dynamics, we tested several natural scenarios to identify uncertainties associated with estimating key population variables. Using a set of eight scenarios, we aimed to diagnose, evaluate, assess performance, and compare the tested models. This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.  Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in structure population.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques. All this data was proceses in @Dornam2021; 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.  To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator **"fleet"** to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.

@@

The process of modeling and statistical analysis of a database can be structured according to the following guidelines:

- Contextualization of the problem. Definition of objectives and variables.
- Experiment design and data collection.
- Recording and preliminary processing of available information.
- Graphical inspection and trend identification.
- Consideration of distributional and relational hypotheses. Proposal of modeling.
- Model adjustment. Comparison and selection of the best model.
- Diagnosis and validation of the adjusted model.
- Assessment of the predictive capacity of the model and prediction.
- Interpretation and conclusions.

## Population Dynamics Model

@@

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population. This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.

Life history arameters taken as priors taken from @Smith2023a and @Maschette2020

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Despite this, integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.

Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP.

The stock assessment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21). SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*. SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data. The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013: @methot2020stock]. The analysis of results and outputs uses R tools and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].

By integrating data from multiple sources and considering spatial heterogeneity, the assessment methodology using Stock Synthesis v.3.30.21 provides a robust framework for evaluating the status of krill populations in the Antarctic Peninsula region. This information is essential for supporting management decisions aimed at ensuring the sustainable use of krill resources in this ecologically sensitive area.

## Environmental factor in modelation

Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

## Predator components in modelation

Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"*. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock].

## Scenarios

To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model  regardng this configurations. In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |

: Scenarios to modelling dynamics in krill in spatial implicit context

These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment. all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

@@

Step to do a good practice in model diagnosis is;

- 1. Convergence. Final convergence criteria is 1.0e-04

@@

## Predator fleet with RW

Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.

Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter. In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process. Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.

Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time. However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.

@@

## Verosimilitud

@@

\newpage
# DISCUSION

@@

@@

```{r mapa, out.width='100%', fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

@@

```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
knitr::include_graphics('Figs/pathmod.png')
```

@@

```{r comparision, fig.cap="Data series using in s2 scenario"}
par(mfrow=c(2,2))
knitr::include_graphics(c('compare8_Fvalue_uncertainty.png',
                          'compare2_spawnbio_uncertainty.png',
                          'compare4_Bratio_uncertainty.png',
                          'compare8_Fvalue_uncertainty.png'))
```

>>>>

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
dir1<-here("s1")
```

@@

The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a stock assessment model. Being able to identify index data relative, systematize, integrate and estimate krill population variables in a context of change and fisheries management is vital for decision making. On the other hand, spatial consideration and ecosystem aspect are another vital component for the application of this kind of approach. To do this, we use Stock Synthesis (SS3) as a statistical platform, a modeling tool widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. This approach has three main objectives. Firstly, integrate fishing, environmental and ecological variables in an integrated model. On the other hand, consider the spatial heterogeneity of the krill population structure. Finally, evaluate the impact of biological and population assumptions on the performance of this type of integrated stock assessment models. The work approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions, which in turn, allows the adequate configuration of control rules that may be useful in the fisheries management applied by CCAMLR of this important resource in the Southern Ocean. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

*Keywords: Krill populations, dynamic population, integrated model, SS3, spatial implicit, ecosystem consideration, management, CCAMLR.*

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013]. This regional organization has tried to conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. There is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables in krill. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others. In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery biological and ecosystem history on krill population over the past 20 years, as well as environmental variables, and predator abundance.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.  Each stratum may have distinct characteristics that affect krill biology (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows you to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques. All this data was proceses in @Dornam2021 and length compositions (Suppl. Mat. 3); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics on krill in the WAP area.  To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell) (Suppl. Mat 3). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series (Figure \@ref(fig:peng)), we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment. Al this data and temporal scale is showed in Figure \@ref(fig:datas2).

### Environmental variable in modelling

Furthermore, this integrated model will take into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

### Predator components in modelling

This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator  abundance and feeding ecology, our model will provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"* how  we describe previously. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.

@@

A conceptual model for krill in the Antarctic Peninsula includes five spatial strata. Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)). Ecological and environmental variables include krill biomass, predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The analysis of results and outputs uses R tools (@R-base) and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].

## Population dynamics Model

@@

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population. This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment. More detail in main equation and model configuration can be found in @methot2020stock. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).

## Stock Assessment Scenarios

To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding this configurations. In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate 48.1                |
|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s2    | "s1" without S-R relation                                                            |
|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
|    s6    | "s1" Ricker S-R relation estimated                                                   |
|    s7    | "s1" w/ set of parameters estimated in @SAM-2024/32                                   |

: Scenarios to modelling dynamics in krill

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @WP2024/34

These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment. All this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

@@

All models tested was the final gradient of the  model was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive definite which shows that the performance is within what was expected for this type of analysis.

Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there is variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall).

Overall, the model's performance is robust, but there is room for improvement in capturing the observed variability in certain fleets and years. In residuals components for `s1`, `s01` and `s7`, the model collects the modes and proportions for each source of information that contains lengths, however, it is the fleet of predators that fails to replicate the data of juvenile individuals (Figure \@ref(fig:fitcom)).

@@

Main variables and forecasting can be found in Table \@ref(tab:mainvar)

@@

@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern.

## Likelihood profile in spatial Model `s1`

The R0 likelihood component profiles for the SMA and HAKE models (Fig. 6) were developed using the function r4ss:SS_profile() (see https://github.com/jabbamodel/ss3diags for example R code). For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources. The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro). The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.

## Cross correlation

@@

Regarding to autocorrelation in krill recruitment in time series analized by scenarios, **S01** most ACF values fall within the significance limits, suggesting no significant autocorrelation, which indicates the series might be random or lack temporal dependency. In **S1**, several ACF values at the initial lags are significant, indicating the presence of autocorrelation in the early periods and suggesting temporal dependency. Similarly, **S2** shows significant autocorrelation in the initial lags, pointing to early temporal dependency as well. **S3** mostly stays within the significance bounds, with some exceptions at the initial lags, hinting at a slight autocorrelation in the early periods, but it is not very pronounced. **S4** mirrors S01, with ACF values within the significance limits, indicating no significant autocorrelation and suggesting randomness in the series. **S5** displays some significant autocorrelation in the initial lags, although less pronounced than in S1 and S2, suggesting a mild early temporal dependency. **S6** follows the pattern of S01 and S4, with no significant autocorrelation, implying a potentially random series. Lastly, **S7** shows ACF values mostly within the significance bounds, with a few exceptions at the initial lags, indicating a slight autocorrelation in the early periods, though not strongly pronounced.In summary, S01, S4, and S6 exhibit no significant autocorrelation, suggesting randomness.

# DISCUSION

Beyond integrating available information for assessing krill population dynamics, we tested several scenarios to identify uncertainties associated with estimating key population variables that aimed to diagnose, evaluate, assess performance, and compare the tested models. This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.

@@

- The main challenge has been to identify sources of information about krill population dynamic. It was possible to integrate the various pieces of information (fishery, survey, environmental, predator, LH
parameters) that are available for krill in 48.1 into an integrated model (SS3) that takes care of the fishing and biological history of this species.
- Conducting an stock assessment and get “best fit” is not the ultimate goal; rather, it is to identify sources of uncertainty associated with krill population dynamics and test their impact on main variables and transfer this type of information into fisheries management in CCAMLR.
- Different scenarios about spatial heterogeneity and the krill stock-recruit relationship (steepness) were tested.
- There are various ways to check the performance of the models. More data sources, more parameters!
- The retrospective pattern of the aggregate model s1 is greater than s01, but the RMSE indicates that model
s01 has less predictive power

# SUPPLEMENTARY MATERIAL

- Supplementary Material 3:  Length composition templates preparation to SS3 from AMLR data [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)

- Supplementary Material 4:  Correlation analysis between. biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)

@@

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```
\begin{landscape}
```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
knitr::include_graphics('Figs/pathmod.png')
```

```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

# leo archivos para plotear y hacer tablas
start1 <- SS_readstarter(file = file.path(dir1,
                                          "starter.ss"),
                              verbose = FALSE)
# note the data and control file names can vary, so are determined from the 
# starter file.
dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
                        verbose = FALSE)
# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
# assumes the correct data structure
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()

parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")

 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

parInit %>%
  kbl(booktabs = TRUE,
      format = "latex",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped"),
                full_width = FALSE,
                font_size=9)%>% 
  pack_rows(index = c("Natural Mortality" = 1,
                        "Growth"= 5,
                        "Length-Weigth Relation" = 2,
                        "Maturity"=2,
                        "Stock-Recruit Relation"=5,
                        "Catchability"=2,
                        "Selectivity"=4))
```

@@

```{r peng,  out.width='100%', fig.cap="Primary penguin species time series abundant populations in colonies located at 48.1"}
knitr::include_graphics('Figs/Cunt_Penguins.png')
```
\end{landscape}

```{r datas2, out.width='70%', fig.cap="Data presence by year for each fleet and data type used in s1 scenario"}
knitr::include_graphics('Figs/data_plot_s2.png')
```

```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
```

```{r res1a, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
```

```{r fitcom, out.width='50%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
par(mfrow=c(3,1))
knitr::include_graphics(c('s01/plots/comp_lenfit__aggregated_across_time.png',
                          's1/plots/comp_lenfit__aggregated_across_time.png',
                          's7/plots/comp_lenfit__aggregated_across_time.png'))

```

```{r resall, out.width='20%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes."}
par(mfrow=c(2,5))
knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))

```
\end{landscape}

```{r likepro, out.width='60%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
knitr::include_graphics('Figs/Profile_s1.png')
```

\begin{landscape}
```{r acf, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-31-1.jpeg')
```
\end{landscape}

\begin{landscape}
```{r recdev, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-30-1.jpeg')
```

@@

```{r comparision, out.width='40%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
par(mfrow=c(2,2))
knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
                          'Figs/compare2_spawnbio_uncertainty.png',
                          'Figs/compare4_Bratio_uncertainty.png',
                          'Figs/compare17_densities_SSB_Virgin.png'))
```
